A comparison of simple models for urban morphogenesis

### 2020-10-01 Docs/Papers/MesoBenchmark/arxiv/v1/paper.tex
submission_mesobenchmark

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig1.png}
	\caption{{\bf .}\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig2.png}
	\caption{{\bf .}\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%

@@

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig3.png}
	\caption{{\bf .}\label{fig:fig3}}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig4.png}
	\caption{{\bf .}\label{fig:fig4}}
\end{figure}
%%%%%%%%%%%%%%

@@

@@

% dynamical calib
\cite{raimbault:halshs-02406539}

@@

@@

\bibliography{biblio}

>>>>

% Use the PLoS provided BiBTeX style
%\bibliographystyle{plos2015}

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{Fig1.png}
	\caption{{\bf Maps of urban morphology indicators for worldwide urban areas.} (Top) Rank-size hierarchy; (Bottom) Moran spatial autocorrelation index. Area of circles gives population.\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{Fig2.png}
	\caption{{\bf Urban form typology. Using an unsupervised k-means clustering algorithm, we find clusters among urban areas.} (Top) Map of cluster belonging; (Bottom) Statistical distribution of indicator values, within each cluster and in time for all dates in the GHSL dataset.\label{fig:fig2}}
\end{figure}
%%%%%%%%%%%%%%

We compare model simulation results to urban form measures computed on worldwide urban areas. We use therefore the Global Human Settlement Layer database (GHSL), which provides an exhaustive worldwide population raster with a 1km resolution \cite{melchiorri2018unveiling}. \cite{Raimbault_2020} has shown the relevance of using this database for worldwide simulation models. It is available at four dates from 1975 onwards, but as not all models are dynamical we use the most recent population configuration only (2015). The database provides a layer of urban areas, within which we extract for the 1000 largest areas in terms of population a covering window ($\pm 25\%$ of extent on each side) from the population raster (note that some windows may be overlapping, as in the case of Hong-Kong which is separated from the main cluster of the Pearl River Delta mega-city region). Morphological indicators are computed on these extracted areas.

We show in Fig.~\ref{fig:fig1} maps of indicators. More particularly, we map the rank-size hierarchy and Moran spatial autocorrelation which have meaningful geographical variations. Rank-size hierarchy will tell if the metropolitan area is strongly dominated by one center or if it more balanced. We retrieve the fact that in Europe, Paris and London are known for such a strong monocentricity, compared to cities in Germany for example. Similarly, mega-city regions in East Asia (Pearl River delta, Yangtze River delta, Beijing-Tianjin) are more polycentric and thus balanced than Wuhan or Seoul. Regarding spatial autocorrelation, we also observe a strong variation in East Asia (Tokyo compared to Chinese megacities for example), and within India (agglomerations in the Gange plain making a cluster of non-correlated, thus highly sprawled areas).

The areas can be clustered following a non-supervised approach. We proceed to a k-means clustering on normalized indicator values, and find $k=4$ clusters as meaningful regarding the derivative of within-cluster variance. A map of cluster belonging and cluster profiles are shown in Fig.~\ref{fig:fig2}. We retrieve the variation within East Asia (Tokyo, Seoul and Shanghai being each in a different cluster) but less in Europe. The main cluster (light blue) corresponds to strongly monocentric urban areas. We see in density distributions of indicator values that clusters have clearly different profiles, which correspond to different typologies of urban morphology that the models try to approximate. For example, cluster 3 (light blue) and 1 (red) have the same level of hierarchy, but the latter has a much higher autocorrelation and entropy, and corresponds thus to more polycentric configurations.

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{Fig3.png}
	\caption{{\bf Examples of generated urban forms for a square world of width $W = 50$.} From Left to Right and Top to Bottom: (i) Correlated percolation model for $r_0^{(C)} = 50$, $\alpha^{(C)} = 0.4$, $n^{(C)} = 3$; (ii) Exponential mixture model for $n^{(E)} = 10$, $r_0^{(E)} = 40$, $\alpha^{(E)} = 1$; (iii) Gravity model for $g^{(G)} = 0.3$, $\gamma_D^{(G)} = 2.5$, $\gamma_P^{(G)} = 0.5$, $P_0^{(G)} = 3$, $P_M^{(G)} = 20000$; (iv) Reaction-diffusion model for $\alpha^{(R)} = 0.8$, $\beta^{(R)} = 0.2$, $d^{(R)} = 1$, $N^{(R)} = 100$, $P_M^{(R)} = 5000$.\label{fig:fig3}}
\end{figure}
%%%%%%%%%%%%%%

@@

We show in Fig.~\ref{fig:fig3} examples of generated urban forms for each model included in this study. Visually, these urban form look rather different. To what extent they are statistically distant for the morphology indicators can only be determined by systematic experiments. The gravity and percolation results look similar, although at a slightly different scale. This particular configuration of the reaction-diffusion model corresponds more to a rural or peri-urban configuration, while the exponential mixture is fuzzy and would resemble a blurred polycentric urban configuration.

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{Fig4.png}
	\caption{{\bf Feasible morphological space for each simulation model.} The lower part of the plot matrix gives a scatterplot between each indicator. We also plot real urban areas in purple. The diagonal gives statistical distribution, while the correlation matrix is given in the upper part, conditionally to each model.\label{fig:fig4}}
\end{figure}
%%%%%%%%%%%%%%

@@

We now turn to the main experiment of this paper: using a diversity search algorithm to determine the full feasible morphological space for each model. Therefore, we use the Pattern Space Exploration (PSE) algorithm \cite{cherel2015beyond} embedded in OpenMOLE. Diversity search was introduced in the field of artificial life as genetic algorithms with the aim to maximize diversity of the population \cite{lehman2008exploiting}. In the case of the PSE algorithm, a novelty criteria leads the search towards new regions of the indicator space, and results are stored in a hitmap. When running long enough, convergence in terms of number of solutions found is generally reached, and one can consider the final population as the feasible space of the model. We run here the algorithm up to 100,000 generations for each model, with grid of step 0.05 in the indicator space. Convergence was reached separately for each model.

We show in Fig~\ref{fig:fig4} a scatterplot of the final population in the morphological space. First of all, we observe that correlation are very different for each model (for example opposite value for gravity and reaction-diffusion between entropy and slope), confirming that the way urban form is produced is very different although the final form may be the same. Then, we observe that the point clouds slightly intersect but also have their own proper space that which no other model can reach. For example, in the entropy-slope space, the reaction-diffusion model is very flexible and covers a large part of space, while gravity is restricted to a small region and correlated percolation and exponential mixture exhibit much less variability with a strong correlation between the two dimensions. The models cover a similar region for the Moran-entropy space, but all fail to capture an important part of real points (in purple), corresponding to configurations with a high Moran index but a relatively low entropy. Models systematically associate a high moran with a high entropy. In other dimensions, real points are reasonably covered. Thus, we can conclude that (i) the different models are complementary in terms of urban forms produced and (ii) that a part of real configurations are approached by some of the models, but other can not be reproduced.

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\begin{center}
	\includegraphics[width=0.6\linewidth]{Fig5.png}
	\end{center}
	\caption{{\bf Overlap between point cloud hypervolumes.}\label{fig:fig5}}
\end{figure}
%%%%%%%%%%%%%%

To better quantify how each model are similar and how they approach real configurations, we compute the hypervolume (in the four dimensional indicator space) corresponding to each point cloud and the intersections between these hypervolumes. We use the \texttt{hypervolume} \texttt{R} package \cite{hypervolume} with a gaussian kernel density estimation with adaptable bandwidth. Then, we compute for each couple of model the ratio between the intersection of hypervolumes and the second model. The non-symmetric relative overlap matrix is plotted in Fig.~\ref{fig:fig5} (diagonal was removed for a better readability). We find that the closer model are the correlated percolation and reaction diffusion, with close to 50\% overlap. Then the gravity point cloud is mostly contained by reaction-diffusion and correlated percolation, but not the other way around. Thus models produce similar points in the parameter space, but most of their output cloud is original. Finally, we find regarding the proximity to real configurations that the best model is the exponential mixture. At the price of producing blurry urban forms, the flexibility therein is the best to approach the best existing configurations. Then comes the reaction-diffusion model, and the least flexible to approach real values is the correlated percolation model.

@@

We have shown that different simple urban morphogenesis models are strongly complementary in the morphological space, confirming that a complementarity of processes also leads to a complementarity in patterns produced. This result rejoins the results obtained by \cite{raimbault2018multi} in the case of transportation network generative models at a similar mesoscopic scale, the results of \cite{raimbault2019generating} for generating building configurations, and the results of \cite{raimbault2018systematic} in the case of dynamical models for systems of cities. Both found a complementarity of approach including different types of processes. We argue that this is further evidence of the multi-dimensionality of urban systems and for a necessity of a plurality of urban models to capture both diverse processes but also outcomes.

@@

% dynamical calib

This work is a first step towards a systematic benchmark of simple urban morphogenesis models. Future work should include the search for an explanation of the unreachable real urban configuration, and possibly alternative models approaching these. Other urban form indicators should also be tested. Finally, dynamical calibration of models remains an open question, investigated in the case of the reaction-diffusion model by \cite{raimbault:halshs-02406539}. An issue is that some models such as the correlated percolation model are not dynamical and should thus be adapted to be calibrated between successive points in time.

@@

We have implemented and systematically compared four very different simple models for urban morphogenesis, including reaction-diffusion, correlated percolation, exponential mixture and a gravitational aggregation model. We applied a diversity search algorithm to obtain the feasible morphological space. The results confirm a complementarity between the models and the relevance of a plurality in urban modeling approaches.

@@

\begin{thebibliography}{10}

\bibitem{batty2018inventing}
Batty M.
\newblock Inventing future cities.
\newblock MIT press; 2018.

\bibitem{williams2000achieving}
Williams K, Burton E, Jenks M.
\newblock Achieving sustainable urban form: an introduction.
\newblock Achieving sustainable urban form. 2000;2000:1--5.

\bibitem{le2012urban}
Le~N{\'e}chet F.
\newblock Urban spatial structure, daily mobility and energy consumption: a
  study of 34 european cities.
\newblock Cybergeo: European Journal of Geography. 2012;.

\bibitem{viguie2012trade}
Vigui{\'e} V, Hallegatte S.
\newblock Trade-offs and synergies in urban climate policies.
\newblock Nature Climate Change. 2012;2(5):334--337.

\bibitem{wegener2004land}
Wegener M, F{\"u}rst F.
\newblock Land-use transport interaction: state of the art.
\newblock Available at SSRN 1434678. 2004;.

\bibitem{milton2019accelerating}
Milton R, Roumpani F.
\newblock Accelerating Urban Modelling Algorithms with Artificial Intelligence.
\newblock In: Proceedings of the 5th International Conference on Geographical
  Information Systems Theory, Applications and Management. vol.~1. INSTICC;
  2019. p. 105--116.

\bibitem{batty1997cellular}
Batty M.
\newblock Cellular automata and urban form: a primer.
\newblock Journal of the American Planning Association. 1997;63(2):266--274.

\bibitem{makse1998modeling}
Makse HA, Andrade JS, Batty M, Havlin S, Stanley HE, et~al.
\newblock Modeling urban growth patterns with correlated percolation.
\newblock Physical Review E. 1998;58(6):7054.

\bibitem{batty2007cities}
Batty M.
\newblock Cities and complexity: understanding cities with cellular automata,
  agent-based models, and fractals.
\newblock The MIT press; 2007.

\bibitem{murcio2015urban}
Murcio R, Morphet R, Gershenson C, Batty M.
\newblock Urban transfer entropy across scales.
\newblock PLoS One. 2015;10(7):e0133780.

\bibitem{batty1989urban}
Batty M, Longley P, Fotheringham S.
\newblock Urban growth and form: scaling, fractal geometry, and
  diffusion-limited aggregation.
\newblock Environment and planning A. 1989;21(11):1447--1472.

\bibitem{murcio2009colored}
Murcio R, Rodr{\'\i}guez-Romo S.
\newblock Colored diffusion-limited aggregation for urban migration.
\newblock Physica A: Statistical Mechanics and its Applications.
  2009;388(13):2689--2698.

\bibitem{murcio2013second}
Murcio R, Sosa-Herrera A, Rodriguez-Romo S.
\newblock Second-order metropolitan urban phase transitions.
\newblock Chaos, Solitons \& Fractals. 2013;48:22--31.

\bibitem{10.1371/journal.pone.0203516}
Raimbault J.
\newblock Calibration of a density-based model of urban morphogenesis.
\newblock PLOS ONE. 2018;13(9):1--18.
\newblock doi:{10.1371/journal.pone.0203516}.

\bibitem{li2019singularity}
Li Y, Rybski D, Kropp JP.
\newblock Singularity cities.
\newblock Environment and Planning B: Urban Analytics and City Science. 2019;
  p. 2399808319843534.

\bibitem{pumain2020conclusion}
Pumain D, Raimbault J.
\newblock Conclusion: Perspectives on urban theories.
\newblock In: Theories and Models of Urbanization. Springer; 2020. p. 303--330.

\bibitem{fotheringham1989spatial}
Fotheringham AS, O'Kelly ME.
\newblock Spatial interaction models: formulations and applications. vol.~1.
\newblock Kluwer Academic Publishers Dordrecht; 1989.

\bibitem{rybski2013distance}
Rybski D, Ros AGC, Kropp JP.
\newblock Distance-weighted city growth.
\newblock Physical Review E. 2013;87(4):042114.

\bibitem{turing1990chemical}
Turing AM.
\newblock The chemical basis of morphogenesis.
\newblock Bulletin of mathematical biology. 1990;52(1-2):153--197.

\bibitem{raimbault2018co}
Raimbault J.
\newblock Co-evolution and morphogenetic systems.
\newblock arXiv preprint arXiv:180311457. 2018;.

\bibitem{bonin2014modelisation}
Bonin O, Hubert JP.
\newblock Mod{\'e}lisation morphog{\'e}n{\'e}tique de moyen terme des villes:
  une sch{\'e}matisation du mod{\`e}le th{\'e}orique de Ritchot et Desmarais
  dans le cadre du mod{\`e}le standard de l'{\'e}conomie urbaine.
\newblock Revue dEconomie Regionale Urbaine. 2014;(3):471--497.

\bibitem{parish2001procedural}
Parish YI, M{\"u}ller P.
\newblock Procedural modeling of cities.
\newblock In: Proceedings of the 28th annual conference on Computer graphics
  and interactive techniques; 2001. p. 301--308.

\bibitem{makse1996method}
Makse HA, Havlin S, Schwartz M, Stanley HE.
\newblock Method for generating long-range correlations for large systems.
\newblock Physical Review E. 1996;53(5):5445.

\bibitem{anas1998urban}
Anas A, Arnott R, Small KA.
\newblock Urban spatial structure.
\newblock Journal of economic literature. 1998;36(3):1426--1464.

\bibitem{zhang2005metrics}
Zhang M, Kukadia N.
\newblock Metrics of urban form and the modifiable areal unit problem.
\newblock Transportation Research Record. 2005;1902(1):71--79.

\bibitem{raimbault2019generating}
Raimbault J, Perret J.
\newblock Generating urban morphologies at large scales.
\newblock In: Artificial Life Conference Proceedings. MIT Press; 2019. p.
  179--186.

\bibitem{10.1371/journal.pone.0225734}
Bosch M.
\newblock PyLandStats: An open-source Pythonic library to compute landscape
  metrics.
\newblock PLOS ONE. 2019;14(12):1--19.
\newblock doi:{10.1371/journal.pone.0225734}.

\bibitem{tsai2005quantifying}
Tsai YH.
\newblock Quantifying urban form: compactness versus' sprawl'.
\newblock Urban studies. 2005;42(1):141--161.

\bibitem{chen2011derivation}
Chen Y.
\newblock Derivation of the functional relations between fractal dimension of
  and shape indices of urban form.
\newblock Computers, Environment and Urban Systems. 2011;35(6):442--451.

\bibitem{schwarz2010urban}
Schwarz N.
\newblock Urban form revisited—Selecting indicators for characterising
  European cities.
\newblock Landscape and urban planning. 2010;96(1):29--47.

\bibitem{raimbault2020scala}
Raimbault J, Perret J, Reuillon R.
\newblock A scala library for spatial sensitivity analysis.
\newblock arXiv preprint arXiv:200710667. 2020;.

\bibitem{reuillon2013openmole}
Reuillon R, Leclaire M, Rey-Coyrehourcq S.
\newblock OpenMOLE, a workflow engine specifically tailored for the distributed
  exploration of simulation models.
\newblock Future Generation Computer Systems. 2013;29(8):1981--1990.

\bibitem{melchiorri2018unveiling}
Melchiorri M, Florczyk AJ, Freire S, Schiavina M, Pesaresi M, Kemper T.
\newblock Unveiling 25 years of planetary urbanization with remote sensing:
  Perspectives from the global human settlement layer.
\newblock Remote Sensing. 2018;10(5):768.

\bibitem{Raimbault_2020}
Raimbault J, Denis E, Pumain D.
\newblock Empowering Urban Governance through Urban Science: Multi-Scale
  Dynamics of Urban Systems Worldwide.
\newblock Sustainability. 2020;12(15):5954.
\newblock doi:{10.3390/su12155954}.

\bibitem{cherel2015beyond}
Ch{\'e}rel G, Cottineau C, Reuillon R.
\newblock Beyond corroboration: Strengthening model validation by looking for
  unexpected patterns.
\newblock PloS one. 2015;10(9):e0138212.

\bibitem{lehman2008exploiting}
Lehman J, Stanley KO.
\newblock Exploiting open-endedness to solve problems through the search for
  novelty.
\newblock In: ALIFE; 2008. p. 329--336.

\bibitem{hypervolume}
Blonder B, with contributions~from David J~Harris. hypervolume: High
  Dimensional Geometry and Set Operations Using Kernel Density Estimation,
  Support Vector Machines, and Convex Hulls; 2019.
\newblock Available from: \url{https://CRAN.R-project.org/package=hypervolume}.

\bibitem{raimbault2018multi}
Raimbault J.
\newblock Multi-modeling the morphogenesis of transportation networks.
\newblock In: Artificial Life Conference Proceedings. MIT Press; 2018. p.
  382--383.

\bibitem{raimbault2018systematic}
Raimbault J.
\newblock A systematic comparison of interaction models for systems of cities.
\newblock In: Conference on Complex Systems 2018; 2018.

\bibitem{raimbault:halshs-02406539}
Raimbault J.
\newblock {Worldwide estimation of parameters for a simple reaction-diffusion
  model of urban growth}.
\newblock In: {International Land-use Symposium 2019}. Paris, France;
  2019.Available from:
  \url{https://halshs.archives-ouvertes.fr/halshs-02406539}.

\end{thebibliography}

### 2020-10-01 Docs/Papers/MesoBenchmark/paper/paper.tex
submission_mesobenchmark

%\documentclass[10pt,letterpaper]{article}
%\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
\documentclass[10pt]{article}
%\usepackage[margin=2cm]{geometry}

@@

% Remove comment for double spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
%\raggedright
%\setlength{\parindent}{0.5cm}
%\textwidth 5.25in 
%\textheight 8.75in

@@

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
%\pagestyle{fancy}
%\fancyhf{}
%%\setlength{\headheight}{27.023pt}
%%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
%\rfoot{\thepage/\pageref{LastPage}}
%\renewcommand{\headrulewidth}{0pt}
%\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
%\fancyheadoffset[L]{2.25in}
%\fancyfootoffset[L]{2.25in}
%\lfoot{\today}

@@

%\begin{document}
%\vspace*{0.2in}

% Title must be 250 characters or less.
%\begin{flushleft}
%\begin{center}
%{\Large
%\textbf\newline{A comparison of simple models for urban morphogenesis} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
%}
%\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
%\\
\title{A comparison of simple models for urban morphogenesis}
\date{}

\author{
Juste Raimbault\textsuperscript{1,2,3,*}
\bigskip\\
\textsuperscript{1} Center for Advanced Spatial Analysis, University College London
\\
\textsuperscript{2} UPS CNRS 3611 ISC-PIF
\\
\textsuperscript{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
\bigskip\\
* juste.raimbault@polytechnique.edu
}

\begin{document}

\maketitle

%\end{flushleft}
%\end{center}
% Please keep the abstract below 300 words
%\section*{Abstract}
\vspace{1cm}
\begin{abstract}
The spatial distribution of population and activities within urban areas, or urban form at the mesoscopic scale, is the outcome of multiple antagonist processes. We propose in this paper to benchmark different models of urban morphogenesis, to systematically compare the urban forms they can produce. Different types of approaches are included, such as a reaction-diffusion model, a gravity-based model, and correlated percolation. Applying a diversity search algorithm, we estimate the feasible space of each model within a space of urban form indicators, in comparison of empirical values for worldwide urban areas. We find a complementarity of the different types of processes, advocating for a plurality of urban models.
\end{abstract}

%\linenumbers

@@

The models are implemented in \texttt{scala} and integrated into the \texttt{spatialdata} library for spatial sensitivity analysis \cite{raimbault2020scala}. The library is bundled as an OpenMOLE plugin for the numerical experiments. OpenMOLE is an open source software for model exploration and validation \cite{reuillon2013openmole} combining model embedding with state-of-the-art exploration methods (including for example sensitivity analysis, design of experiments, calibration with genetic algorithms) and a transparent distribution of computations on high performance computing environments. In our case, we use its workflow system and an integrated algorithm to determine the feasible space of models.

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig1.png}
	\caption{{\bf .}\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig2.png}
	\caption{{\bf .}\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%

@@

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig3.png}
	\caption{{\bf .}\label{fig:fig3}}
\end{figure}
%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	\includegraphics[width=\linewidth]{figures/Fig4.png}
	\caption{{\bf .}\label{fig:fig4}}
\end{figure}
%%%%%%%%%%%%%%

>>>>

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}
%\documentclass[10pt]{article}
%\usepackage[margin=2cm]{geometry}

@@

% Remove comment for double spacing
\usepackage{setspace} 
\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in 
\textheight 8.75in

@@

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

@@

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{A comparison of simple models for urban morphogenesis} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Juste Raimbault\textsuperscript{1,2,3,*}
\bigskip\\
\textsuperscript{1} Center for Advanced Spatial Analysis, University College London
\\
\textsuperscript{2} UPS CNRS 3611 ISC-PIF
\\
\textsuperscript{3} UMR CNRS 8504 G{\'e}ographie-cit{\'e}s
\bigskip\\
* juste.raimbault@polytechnique.edu

\end{flushleft}

% Please keep the abstract below 300 words
\section*{Abstract}
\vspace{1cm}
The spatial distribution of population and activities within urban areas, or urban form at the mesoscopic scale, is the outcome of multiple antagonist processes. We propose in this paper to benchmark different models of urban morphogenesis, to systematically compare the urban forms they can produce. Different types of approaches are included, such as a reaction-diffusion model, a gravity-based model, and correlated percolation. Applying a diversity search algorithm, we estimate the feasible space of each model within a space of urban form indicators, in comparison of empirical values for worldwide urban areas. We find a complementarity of the different types of processes, advocating for a plurality of urban models.

% urban morphogenesis; reaction-diffusion; correlated percolation; gravity model; novelty search

\linenumbers

@@

The models are implemented in \texttt{scala} and integrated into the \texttt{spatialdata} library for spatial sensitivity analysis \cite{raimbault2020scala}. The library is bundled as an OpenMOLE plugin for the numerical experiments. OpenMOLE is an open source software for model exploration and validation \cite{reuillon2013openmole} combining model embedding with state-of-the-art exploration methods (including for example sensitivity analysis, design of experiments, calibration with genetic algorithms) and a transparent distribution of computations on high performance computing environments. In our case, we use its workflow system and an integrated algorithm to determine the feasible space of models. Source code for models is available in the open source code of the \texttt{spatialdata} library at \url{https://github.com/openmole/spatialdata}. Source code for numerical experiments, simulation data and results used in this paper are available on an open git repository at \url{https://github.com/JusteRaimbault/UrbanGrowth}.

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	%\includegraphics[width=\linewidth]{figures/Fig1.png}
	\vspace{2cm}
	\caption{{\bf Maps of urban morphology indicators for worldwide urban areas.} (Top) Rank-size hierarchy; (Bottom) Moran spatial autocorrelation index. Area of circles gives population.\label{fig:fig1}}
\end{figure}
%%%%%%%%%%%%%%

%%%%%%%%%%%%%%
\begin{figure}[!h]
	%\includegraphics[width=\linewidth]{figures/Fig2.png}
	\vspace{2cm}
	\caption{{\bf Urban form typology. Using an unsupervised k-means clustering algorithm, we find clusters among urban areas.} (Top) Map of cluster belonging; (Bottom) Statistical distribution of indicator values, within each cluster and in time for all dates in the GHSL dataset.\label{fig:fig2}}
\end{figure}
%%%%%%%%%%%%%%

We compare model simulation results to urban form measures computed on worldwide urban areas. We use therefore the Global Human Settlement Layer database (GHSL), which provides an exhaustive worldwide population raster with a 1km resolution \cite{melchiorri2018unveiling}. \cite{Raimbault_2020} has shown the relevance of using this database for worldwide simulation models. It is available at four dates from 1975 onwards, but as not all models are dynamical we use the most recent population configuration only (2015). The database provides a layer of urban areas, within which we extract for the 1000 largest areas in terms of population a covering window ($\pm 25\%$ of extent on each side) from the population raster (note that some windows may be overlapping, as in the case of Hong-Kong which is separated from the main cluster of the Pearl River Delta mega-city region). Morphological indicators are computed on these extracted areas.

We show in Fig.~\ref{fig:fig1} maps of indicators. More particularly, we map the rank-size hierarchy and Moran spatial autocorrelation which have meaningful geographical variations. Rank-size hierarchy will tell if the metropolitan area is strongly dominated by one center or if it more balanced. We retrieve the fact that in Europe, Paris and London are known for such a strong monocentricity, compared to cities in Germany for example. Similarly, mega-city regions in East Asia (Pearl River delta, Yangtze River delta, Beijing-Tianjin) are more polycentric and thus balanced than Wuhan or Seoul. Regarding spatial autocorrelation, we also observe a strong variation in East Asia (Tokyo compared to Chinese megacities for example), and within India (agglomerations in the Gange plain making a cluster of non-correlated, thus highly sprawled areas).

The areas can be clustered following a non-supervised approach. We proceed to a k-means clustering on normalized indicator values, and find $k=4$ clusters as meaningful regarding the derivative of within-cluster variance. A map of cluster belonging and cluster profiles are shown in Fig.~\ref{fig:fig2}. We retrieve the variation within East Asia (Tokyo, Seoul and Shanghai being each in a different cluster) but less in Europe. The main cluster (light blue) corresponds to strongly monocentric urban areas. We see in density distributions of indicator values that clusters have clearly different profiles, which correspond to different typologies of urban morphology that the models try to approximate. For example, cluster 3 (light blue) and 1 (red) have the same level of hierarchy, but the latter has a much higher autocorrelation and entropy, and corresponds thus to more polycentric configurations.

@@

%%%%%%%%%%%%%%
\begin{figure}[!h]
	%\includegraphics[width=\linewidth]{figures/Fig3.png}
	\vspace{2cm}
	\caption{{\bf Examples of generated urban forms for a square world of width $W = 50$.} From Left to Right and Top to Bottom: (i) Correlated percolation model for $r_0^{(C)} = 50$, $\alpha^{(C)} = 0.4$, $n^{(C)} = 3$; (ii) Exponential mixture model for $n^{(E)} = 10$, $r_0^{(E)} = 40$, $\alpha^{(E)} = 1$; (iii) Gravity model for $g^{(G)} = 0.3$, $\gamma_D^{(G)} = 2.5$, $\gamma_P^{(G)} = 0.5$, $P_0^{(G)} = 3$, $P_M^{(G)} = 20000$; (iv) Reaction-diffusion model for $\alpha^{(R)} = 0.8$, $\beta^{(R)} = 0.2$, $d^{(R)} = 1$, $N^{(R)} = 100$, $P_M^{(R)} = 5000$.\label{fig:fig3}}
\end{figure}
%%%%%%%%%%%%%%

@@

We show in Fig.~\ref{fig:fig3} examples of generated urban forms for each model included in this study. Visually, these urban form look rather different. To what extent they are statistically distant for the morphology indicators can only be determined by systematic experiments. The gravity and percolation results look similar, although at a slightly different scale. This particular configuration of the reaction-diffusion model corresponds more to a rural or peri-urban configuration, while the exponential mixture is fuzzy and would resemble a blurred polycentric urban configuration.

%%%%%%%%%%%%%%
\begin{figure}[!h]
	%\includegraphics[width=\linewidth]{figures/Fig4.png}
	\vspace{2cm}
	\caption{{\bf Feasible morphological space for each simulation model.} The lower part of the plot matrix gives a scatterplot between each indicator. We also plot real urban areas in purple. The diagonal gives statistical distribution, while the correlation matrix is given in the upper part, conditionally to each model.\label{fig:fig4}}
\end{figure}
%%%%%%%%%%%%%%

@@

We now turn to the main experiment of this paper: using a diversity search algorithm to determine the full feasible morphological space for each model. Therefore, we use the Pattern Space Exploration (PSE) algorithm \cite{cherel2015beyond} embedded in OpenMOLE. Diversity search was introduced in the field of artificial life as genetic algorithms with the aim to maximize diversity of the population \cite{lehman2008exploiting}. In the case of the PSE algorithm, a novelty criteria leads the search towards new regions of the indicator space, and results are stored in a hitmap. When running long enough, convergence in terms of number of solutions found is generally reached, and one can consider the final population as the feasible space of the model. We run here the algorithm up to 100,000 generations for each model, with grid of step 0.05 in the indicator space. Convergence was reached separately for each model.

We show in Fig~\ref{fig:fig4} a scatterplot of the final population in the morphological space. First of all, we observe that correlation are very different for each model (for example opposite value for gravity and reaction-diffusion between entropy and slope), confirming that the way urban form is produced is very different although the final form may be the same. Then, we observe that the point clouds slightly intersect but also have their own proper space that which no other model can reach. For example, in the entropy-slope space, the reaction-diffusion model is very flexible and covers a large part of space, while gravity is restricted to a small region and correlated percolation and exponential mixture exhibit much less variability with a strong correlation between the two dimensions. The models cover a similar region for the Moran-entropy space, but all fail to capture an important part of real points (in purple), corresponding to configurations with a high Moran index but a relatively low entropy. Models systematically associate a high moran with a high entropy. In other dimensions, real points are reasonably covered. Thus, we can conclude that (i) the different models are complementary in terms of urban forms produced and (ii) that a part of real configurations are approached by some of the models, but other can not be reproduced.

%%%%%%%%%%%%%%
\begin{figure}[!h]
	%\begin{center}
	%\includegraphics[width=0.6\linewidth]{figures/Fig5.png}
	%\end{center}
	\vspace{2cm}
	\caption{{\bf Overlap between point cloud hypervolumes.}\label{fig:fig5}}
\end{figure}
%%%%%%%%%%%%%%

To better quantify how each model are similar and how they approach real configurations, we compute the hypervolume (in the four dimensional indicator space) corresponding to each point cloud and the intersections between these hypervolumes. We use the \texttt{hypervolume} \texttt{R} package \cite{hypervolume} with a gaussian kernel density estimation with adaptable bandwidth. Then, we compute for each couple of model the ratio between the intersection of hypervolumes and the second model. The non-symmetric relative overlap matrix is plotted in Fig.~\ref{fig:fig5} (diagonal was removed for a better readability). We find that the closer model are the correlated percolation and reaction diffusion, with close to 50\% overlap. Then the gravity point cloud is mostly contained by reaction-diffusion and correlated percolation, but not the other way around. Thus models produce similar points in the parameter space, but most of their output cloud is original. Finally, we find regarding the proximity to real configurations that the best model is the exponential mixture. At the price of producing blurry urban forms, the flexibility therein is the best to approach the best existing configurations. Then comes the reaction-diffusion model, and the least flexible to approach real values is the correlated percolation model.
